---
title: "P8105_HW1_deg2163"
author: Diana Ballesteros Gonzalez 
date: 2018-09-16
output: github_document
---

Homework Assignment 1 

#Problem 1
Create a data frame comprised of:

- A random sample of size 10 from a uniform [0,5] distribution.
- A logical vector indicating whether elements of the sample are greater than 2.
- A (length-10) character vector.
- A (length-10) factor vector.

```{r setup, include = FALSE}
library(tidyverse)
```

```{r df_create}

set.seed(123)

## create data frame with four vectors 
problem_one_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5), 
  vec_logical = (vec_numeric > 2),
  vec_char = c("My", "name", "is", "Diana", "Ballesteros", "Gonzalez", "and", "I", "love", "tacos."),
  vec_factor = factor(c("red", "orange", "yellow", "green", "blue", "orange", "purple", "indigo", "black", "grey"))
) 
```

- The mean for the a random sample of size 10from a uniform [0,5] distribution is `r mean(problem_one_df$vec_numeric)`. I was able to get a mean value for this vector because it averaged the 10 random numbers that it reproduced. 

- The mean for a logical vector indicating whether elements of the sample are greater than 2 is `r mean(problem_one_df$vec_logical)`. I was able to get a mean value for this vector because it calculated the proportion of True statements to False statments.

- The mean of a (length-10) character vector is `r mean(problem_one_df$vec_char)` because you cannot take an average of character values. 


- The mean of a (length-10) factor vector is `r mean(problem_one_df$vec_factor)` because you cannot take an average of categorical variables.

# Problem 2
Create a data frame comprised of:

- x: a random sample of size 1000 from a standard Normal distribution
- y: a random sample of size 1000 from a standard Normal distribution
- A logical vector indicating whether the x + y > 0
- A numeric vector created by coercing the above logical vector
- A factor vector created by coercing the above logical vector





