---
title: "P8105_HW1_deg2163"
author: Diana Ballesteros Gonzalez 
date: 2018-09-16
output: github_document
---

Homework Assignment 1 

#Problem 1
Create a data frame comprised of:

- A random sample of size 10 from a uniform [0,5] distribution.
- A logical vector indicating whether elements of the sample are greater than 2.
- A (length-10) character vector.
- A (length-10) factor vector.

```{r setup}
library(tidyverse)

## create data frame with four vectors 
problem_one_df = tibble(
  vec_numeric = runif(10, 0, 5), 
  vec_char = c("My", "name", "is", "Diana", "Ballesteros", "Gonzalez", "and", "I", "love", "tacos."),
  vec_logical = (vec_numeric > 2),
  vec_factor = factor(c("red", "orange", "yellow", "green", "blue", "orange", "purple", "indigo", "black", "grey"))
) 
```

The mean is `r mean(problem_one_df$vec_numeric)`.
The mean is `r mean(problem_one_df$vec_char)`.
The mean is `r mean(problem_one_df$vec_logical)`.
The mean is `r mean(problem_one_df$vec_factor)`.

# Problem 2
Create a data frame comprised of:

- x: a random sample of size 1000 from a standard Normal distribution
- y: a random sample of size 1000 from a standard Normal distribution
- A logical vector indicating whether the x + y > 0
- A numeric vector created by coercing the above logical vector
- A factor vector created by coercing the above logical vector

Here's a **code chunk** that samples from 
a _normal distribution_:

```{r}
samp = rnorm(100)
length(samp)
```





