---
title: "P8105_HW1_deg2163"
author: Diana Ballesteros Gonzalez 
date: 2018-09-20
output: github_document
---

Homework Assignment 1 

#**Problem 1**
Create a data frame comprised of:

* A random sample of size 10 from a uniform [0,5] distribution.
* A logical vector indicating whether elements of the sample are greater than 2.
* A (length-10) character vector.
* A (length-10) factor vector.

```{r setup, include = FALSE}
library(tidyverse)
```

```{r df_create}

## set the seed to ensure reproducibility
set.seed(123)

## create data frame with four vectors 
problem_one_df = tibble(
  vec_numeric = runif(10, 0, 5), 
  vec_logical = (vec_numeric > 2),
  vec_char = c("My", "name", "is", "Diana", "Ballesteros", "Gonzalez", "and", "I", "love", "tacos"),
  vec_factor = factor(c("red", "orange", "yellow", "green", "blue", "orange", "purple", "indigo", "black", "grey"))
) 
```

* The mean for the a random sample of size 10from a uniform [0,5] distribution is `r mean(problem_one_df$vec_numeric)`. I was able to get a mean value for this vector because it averaged the 10 random numbers that it reproduced. 

* The mean for a logical vector indicating whether elements of the sample are greater than 2 is `r mean(problem_one_df$vec_logical)`. I was able to get a mean value for this vector because it calculated the proportion of true statements out of the 10.

* The mean of a (length-10) character vector is `r mean(problem_one_df$vec_char)` because you cannot take an average of character values. 


* The mean of a (length-10) factor vector is `r mean(problem_one_df$vec_factor)` because you cannot take an average of categorical variables.

###Explanation for what happens when your run as.numeric with the logical, character, and factor vectors
```{r change_types, include=FALSE}
##convert logical, character, and factor to numeric

as.numeric(problem_one_df$vec_logical)

as.numeric(problem_one_df$vec_char)

as.numeric(problem_one_df$vec_factor)
```

* For the logical vector, I had not problem because it converted the true and false statements to numbers where, 0 = False and 1 = True. 
* For the character vector, I received a warning message stating "NAs introduced by coercion."
* For the factor vector, you get random numbers assigned to each factor. 

##Converting from Character to Factor to Numeric
```{r convert_char_factor_numeric}
as.factor(problem_one_df$vec_char)
as.numeric(as.factor(problem_one_df$vec_char))
```
For the conversion of a character to factor to numeric, I was ablw to obtain values that assigned a random number to each string of characters. 

##Converting from Character to Factor to Numeric
```{r convert_factor_character_numeric}
as.character(problem_one_df$vec_factor)
as.numeric(as.character(problem_one_df$vec_factor))
```
For the conversion of character to a factor to numeric, I received a warning message because it cannot produce a numeric value for a character vector. 

# **Problem 2**
Create a data frame comprised of:

- x: a random sample of size 1000 from a standard Normal distribution
- y: a random sample of size 1000 from a standard Normal distribution
- A logical vector indicating whether the x + y > 0
- A numeric vector created by coercing the above logical vector
- A factor vector created by coercing the above logical vector





